---
title: "Hatching success: climwin and bayesian regresson"
output: html_notebook
---
```{r}
# ----------------- LIBRARIES ----------------- #
library(MASS)
library(Matrix)
library(lme4)
library(readxl)
library(arm)
library(rmarkdown)
library(ggplot2)
library(MuMIn)
library(climwin)
library(rstanarm)
library(shinystan)
library(emmeans)
library(magrittr)
library(dplyr)
library(sjPlot)
library(sjstats)
library(modelr)
```
```{r}
# ----------------- DATA LOAD ----------------- #
#####Clim windows#####

Clim <- read_excel("C:/Users/.../Dataset.xlsx", sheet = "SST")
View(Clim)

Clim2 <- read_excel("C:/Users/.../Dataset.xlsx", sheet = "Rainfall")

View(Clim2)

Biol  <- read_excel("C:/Users/.../Dataset.xlsx",sheet = "Hprop_clim")
View(Biol)
names(Biol)

#####Bayesian regression#####
data<- read_excel("C:/Users/.../Dataset.xlsx",sheet = "Hprop_bayes")
```


```{r}
# ----------------- VARIABLES ----------------- #
#####Clim windows#####
# RANDOM EFFECTS
RDATE<-as.factor(Biol$MEAN_LAY_Date)
Y<-as.factor(Biol$YEAR)
MID<-as.factor(Biol$MALE_ID)
FID<-as.factor(Biol$FEMALEID)
#Fixed
FS<-Biol$HATC_PROP

#Dependent
BROOD<-as.factor(Biol$CLUTCH_SIZE)
MAGE<-rescale(Biol$AGEMALE)
FAGE<-rescale(Biol$AGEFEM)
PR<-rescale(Biol$JULIAN_LAY_DAY)


#####Bayesian regression#####
# RANDOM EFFECTS

Y<-as.factor(data$YEAR)
RDATE<-as.factor(data$MEAN_LAY_Date)
MID<-as.factor(data$MALE_ID)
FID<-as.factor(data$FEMALEID)

# DEPENDENT VARIABLE
FS<-data$FLEDGED_PROP

# INDEPENDENT VARIABLES
PR<-rescale(data$JULIAN_LAY_DAY)
BROOD<-as.factor(data$CLUTCH_SIZE)
RAIN<-rescale(data$RAINFALL)
SSTz<-rescale(data$SST)


```

```{r}
```{r}
#####Clim windows#####
#Basal model selection
m1<-lmer(HS ~ (1|Y)+(1|FID)+(1|MID)+(1|RDATE)+BROOD+PR,data=Biol)
m2<-lmer(HS ~ (1|Y)+(1|FID)+(1|MID)+(1|RDATE)+BROOD +PR+ FAGE + I(FAGE^2),data=Biol)
m3<-lmer(HS ~ (1|Y)+(1|FID)+(1|MID)+(1|RDATE)+BROOD +PR+ MAGE+I(MAGE^2),data=Biol)
m4<-lmer(HS ~ (1|Y)+(1|FID)+(1|MID)+(1|RDATE)+BROOD +PR+ FAGE+MAGE,data=Biol)
m5<-lmer(HS ~ (1|Y)+(1|FID)+(1|MID)+(1|RDATE)+BROOD +PR+ FAGE + I(FAGE^2)+MAGE+I(MAGE^2),data=Biol)


v<-model.sel(m1,m2,m3,m4,m5)
v
```


```{r}
#Set date format
Clim$Date<- as.Date(Clim$Date,format="%d %b %Y")
Clim2$Date<- as.Date(Clim2$Date,format="%d %b %Y")
Biol$HATCH_DATE<- as.Date(Biol$HATCH_DATE,format="%d %b %Y")

```


```{r}
#Run sliding window analysis
hatch1<-slidingwin(xvar=list(rescale(Clim$SST),rescale(Clim2$RAIN)),exclude = NA,cdate=Clim$Date,bdate=Biol$HATCH_DATE,baseline=glmer(HS~ (1|MID)+(1|FID)+(1|Y)+ (1|DAT)+ PR + N + FAGE+I(FAGE^2),family = binomial("logit"),weights = Biol$CLUTCH_SIZE,control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5))),range= c(6,0),type="relative",
                  cmissing = "method1",
                  stat=c("mean","max","min","sum"),
                  func = c("lin","quad"),cinterval="week")
#Run random windows
hatch1r<-randwin(xvar=list(rescale(Clim$SST),rescale(Clim2$RAIN)),cdate=Clim$Date,bdate=Biol$HATCH_DATE,baseline= glmer(HS~ (1|MID)+(1|FID)+(1|Y)+ (1|DAT)+ PR + N+ FAGE+I(FAGE^2),family = binomial("logit"),weights = Biol$CLUTCH_SIZE,control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5))), range= c(6,0), type="relative",stat=c("mean","max","min","sum"),func = c("lin","quad"),cinterval="week",cmissing = "method1",repeats = 50)

#When running the GLMMs, a singular fit warning is displayed. Both MID and FID random effects are estimated as zero. Both random terms were dropped from the final analysis.As Bolker points out, dropping this terms from the models do not affect any of the estimated quantities. See https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#singular-models-random-effect-variances-estimated-as-zero-or-correlations-estimated-as---1


```

```{r}
#Show most supported windows
hatch1$combos
```


```{r}

#Calculate Pc value for each most supported model
pvalue(dataset = hatch1[[1]]$Dataset, datasetrand = hatch1r[[1]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[2]]$Dataset, datasetrand = hatch1r[[2]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[3]]$Dataset, datasetrand = hatch1r[[3]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[4]]$Dataset, datasetrand = hatch1r[[4]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[5]]$Dataset, datasetrand = hatch1r[[5]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[6]]$Dataset, datasetrand = hatch1r[[6]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[7]]$Dataset, datasetrand = hatch1r[[7]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[8]]$Dataset, datasetrand = hatch1r[[8]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[9]]$Dataset, datasetrand = hatch1r[[9]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[10]]$Dataset, datasetrand = hatch1r[[10]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[11]]$Dataset, datasetrand = hatch1r[[11]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[12]]$Dataset, datasetrand = hatch1r[[12]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[13]]$Dataset, datasetrand = hatch1r[[13]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[14]]$Dataset, datasetrand = hatch1r[[14]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[15]]$Dataset, datasetrand = hatch1r[[15]], metric = "C", sample.size = 26)

pvalue(dataset = hatch1[[16]]$Dataset, datasetrand = hatch1r[[16]], metric = "C", sample.size = 26)
```


```{r}
#Correlation between the SST and rainfall windows
cross1<-crosswin(xvar=list(rescale(Clim$SST)),xvar2 = list(rescale(Clim2$RAIN)),stat ="mean",stat2 = "min",cdate = Clim$Date,bdate = Biol$HATCH_DATE,range = c(6,0),type = "relative",cmissing = "method1",cinterval = "week")

plotcor(cross1,type="C")
```




```{r}
#Inspection of the best supported models
plothist(dataset = hatch1[[9]]$Dataset, datasetrand = hatch1r[[9]])

plothist(dataset = hatch1[[14]]$Dataset, datasetrand = hatch1r[[14]])
```

```{r}
#Extraction of weather conditions for further inspection
Biol$signal1<-hatch1[[9]]$BestModelData$climate
S1<-Biol$signal1


Biol$signal2<-hatch1[[14]]$BestModelData$climate
S2<-Biol$signal2
```
```{r}
#Rerun of climwin analysis: searching rainfall's climatic window values while controlling for sst 

hatch_sst_on_rain<-slidingwin(xvar=list(rescale(Clim2$RAIN)),exclude = NA,
                        cdate=Clim$Date,
                        bdate=Biol$HATCH_DATE,baseline= glmer(HS~ (1|Y)+ (1|DAT)+ PR + N+ FAGE+I(FAGE^2)+S1+I(S1^2),family = binomial("logit"),weights = Biol$CLUTCH_SIZE,control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5))), range= c(6,0),
                        type="relative",
                       cmissing = "method1",
                        stat=c("mean","max","sum","min"),
                        func = c("lin","quad"),
                        cinterval="week")

hatch_sst_on_rainr<-randwin(xvar=list(rescale(Clim2$RAIN)),exclude = NA,
                        cdate=Clim$Date,
                        bdate=Biol$HATCH_DATE,baseline= glmer(HS~ (1|Y)+ (1|DAT)+ PR + N+ FAGE+I(FAGE^2)+S1+I(S1^2),family = binomial("logit"),weights = Biol$CLUTCH_SIZE,control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5))),
                        range= c(6,0),
                        type="relative",
                       cmissing = "method1",
                        stat=c("mean","max","sum","min"),
                        func = c("lin","quad"),
                        cinterval="week",repeats=50)

```

```{r}
#Show most supported windows

hatch_sst_on_rain$combos
```


```{r}
#Calculate Pc value for each most supported model
pvalue(dataset = hatch_sst_on_rain[[1]]$Dataset, datasetrand = hatch_sst_on_rainr[[1]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_sst_on_rain[[2]]$Dataset, datasetrand = hatch_sst_on_rainr[[2]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_sst_on_rain[[3]]$Dataset, datasetrand = hatch_sst_on_rainr[[3]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_sst_on_rain[[4]]$Dataset, datasetrand = hatch_sst_on_rainr[[4]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_sst_on_rain[[5]]$Dataset, datasetrand = hatch_sst_on_rainr[[5]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_sst_on_rain[[6]]$Dataset, datasetrand = hatch_sst_on_rainr[[6]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_sst_on_rain[[7]]$Dataset, datasetrand = hatch_sst_on_rainr[[7]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_sst_on_rain[[8]]$Dataset, datasetrand = hatch_sst_on_rainr[[8]], metric = "C", sample.size = 26)
```


```{r}
#Rerun of climwin analysis: searching sst's climatic window values while controlling for rainfall 
hatch_rain_on_sst<-slidingwin(xvar=list(rescale(Clim$SST)),exclude = NA,
                        cdate=Clim$Date,
                         bdate=Biol$HATCH_DATE,baseline= glmer(HS~ (1|Y)+ (1|DAT)+ PR + N+ FAGE+I(FAGE^2)+S2+I(S2^2),family = binomial("logit"),weights = Biol$CLUTCH_SIZE,control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5))),
                        range= c(6,0),
                        type="relative",
                       cmissing = "method1",
                        stat=c("mean","max","sum","min"),
                        func = c("lin","quad"),
                        cinterval="week")
hatch_rain_on_sstR<-randwin(xvar=list(rescale(Clim$SST)),exclude = NA,
                        cdate=Clim$Date,
                         bdate=Biol$HATCH_DATE,baseline= glmer(HS~ (1|Y)+ (1|DAT)+ PR + N+ FAGE+I(FAGE^2)+S2+I(S2^2),family = binomial("logit"),weights = Biol$CLUTCH_SIZE,control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun=2e5))),
                        range= c(6,0),
                        type="relative",
                       cmissing = "method1",
                        stat=c("mean","max","sum","min"),
                        func = c("lin","quad"),
                        cinterval="week",repeats=50)
```

```{r}
hatch_rain_on_sst$combos
```



```{r}
pvalue(dataset = hatch_rain_on_sst[[1]]$Dataset, datasetrand = hatch_rain_on_sstR[[1]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_rain_on_sst[[2]]$Dataset, datasetrand = hatch_rain_on_sstR[[2]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_rain_on_sst[[3]]$Dataset, datasetrand = hatch_rain_on_sstR[[3]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_rain_on_sst[[4]]$Dataset, datasetrand = hatch_rain_on_sstR[[4]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_rain_on_sst[[5]]$Dataset, datasetrand = hatch_rain_on_sstR[[5]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_rain_on_sst[[6]]$Dataset, datasetrand = hatch_rain_on_sstR[[6]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_rain_on_sst[[7]]$Dataset, datasetrand = hatch_rain_on_sstR[[7]], metric = "C", sample.size = 26)

pvalue(dataset = hatch_rain_on_sst[[8]]$Dataset, datasetrand = hatch_rain_on_sstR[[8]], metric = "C", sample.size = 26)

plothist(dataset = hatch_rain_on_sst[[5]]$Dataset, datasetrand = hatch_rain_on_sstR[[5]])
```


```{r}
library(sjPlot)
library(sjstats)
library(modelr)
```
```{r}
#Extraction of weather values for the best supported climate window
write.csv(hatch_rain_on_sst[[5]]$BestModelData,"hatch_rain_on_sst.csv")
```


```{r}
#####Bayesian regression#####


options(mc.cores = parallel::detectCores())

m1<-stan_glmer(HP~(1|Y)+(1|D)+(1|MID)+(1|FID)+E+SSTz+I(SSTz^2)+RAINz+I(RAINz^2)+PU+FAGE+I(FAGE^2),family = binomial,weights = data$CLUTCH_SIZE,chains=5,cores=8, prior = normal(0,1),iter=10000,seed=1234,data = data)


#Posterior predictive checks
launch_shinystan(m1)

```

```{r}
print(m1)
```

```{r}
tidy(m1,conf.int = TRUE, conf.method = "HPDinterval",conf.level = 0.89)  

```


```{r}
library(emmeans)
```
```{r}
#Estimate median value of each level of a category
m1.1<-ref_grid(m1)
m1.1
regrid(emmeans(m1.1,"E"))%>%hpd.summary(prob=0.89)
```



```{r}
####Coefficient of variation####
data<- read_excel("C:/Users/.../Dataset.xlsx", sheet = "Hprop_CV")
```
```{r}
CVS<-rescale(data$CV_SST)
CVR<-rescale(data$CV_RAIN)
HS<-data$`MEAN_HATCH`

```

```{r}
m1<-stan_glm(HS~CVS+CVR,weights=data$count,family = binomial,chains=5,cores=8, prior = normal(0,1),iter=10000,seed=1234,data = data)
launch_shinystan(m1)
```

```{r}
print(m1)
tidy(m1,conf.int = TRUE, conf.method = "HPDinterval",conf.level = 0.89)  
```



